\documentclass{article}

\usepackage{color}
\usepackage{listings}
\lstset{ %
language=Java,                % the language of the code
basicstyle=\footnotesize,       % the size of the fonts that are used for the code
numbers=left,                   % where to put the line-numbers
numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1, each line 
                                % will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
frame=single,                   % adds a frame around the code
tabsize=4,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
%breaklines=true,                % sets automatic line breaking
breakatwhitespace=false        % sets if automatic breaks should only happen at whitespace
}

\usepackage{fullpage}
\linespread{1.3}

\begin{document}

\title{JEFRi Executive Summary}
\maketitle
\tableofcontents
\newpage
\linespread{1.6}

\section{JEFRi}
JEFRi, the Javascript Entity Framework Runtime, is a suite of technologies to
make development easy. JEFRi provides a comprehensive approach to building
data-centric software programs. JEFRi lets developers focus on what's important
--- the business data, the business logic, and the business' presentation. With
these core pieces, JEFRi also provides an extensible framework that provides
common, off-the-shelf modules to meet many shared business needs, including
authentication, access control, data reporting, internationalization, and much
more.

\section{The Future of Frameworks}
JEFRi builds on a concept of an entity model, a concept that has had growing
support from a variety of established industry players, including Hibernate on
Oracle's Java platform, and Microsoft's .NET Entity Framework. Entity Frameworks
encourage programs to write in the language of the business' problem domain.
Instead of referencing table 3, column 5 against table 6, column 2, entity
frameworks let the developer write code to get a list of all the client's
brands. By being so much more expressive than an access database, JEFRi lets
developers spend time on the hard problems your buiness faces.

\section{Truly Multiplatform}
Further, JEFRi makes it trivial to migrate to another platform or vendor in the
future. Written in Javascript, JEFRi runs in every browser, on every mobile
phone, and in a rapidly growing number of backend platforms. With Microsoft's
announcement of Windows 8's heavy reliance on HTML, Javascript, and CSS to power
the core of the user interface, there is no better time to encourage developers
to pick up Javascript and a solid Javascript framework. The knowledge will serve
developers on nearly every platform available today.

\section{Tools of the Trade}
JEFRi makes use of conceptual Entities, descriptions of things in a business
model. An entity could be a customer in a CRM, or a blog post for a website.
These entities are connected with relationships, describing how all the
business' data works together. The customer will have many contact cards, while
a blog post has a single author. The JEFRi modeling tool makes it easy to
quickly and correctly express a variety of possible descriptions of the
business' data, and automatically creates all the necessary information for
persisting that data in any database system, from a free MySQL installation, a
powerful Oracle database, or on Amazon's S3 cloud services.

\section{Rapid Prototyping}
JEFRi comes with a series of libraries to rapidly prototype working applications
from a JEFRi model. The data binding modules will take a JEFRi model and provide
a complete web application, including pages, forms, and dynamic requests,
automatically. This code itself is well-formatted and easy to style as is, but
is also completey customizable to fit any design or workflow necessary. This
lets developers focus their attention on areas needing special work, while
levraging existing code to fill in non-critical sections of the application.
With the JEFRi modeler and JEFRi binding, developers can create a
fully-functional, customized CRM system in under a week.

\section{Mobile Powered}
Because modern Android and Apple devices utilize fully-featured web browsers,
JEFRi applications can be run on mobile devices with no change to the
application's code. If an application needs to leverage more features of the
mobile platform, JEFRi makes it easy to replace only the pieces that change with
the mobile platform, primarily the layout and design of the user interface.
Beyond replacing mobile-specific code, JEFRi can make use of a variety of
underlying technologies and aggressive caching to transparently move on and off
data networks, while using the application to its full functionality. This can
be very handy when a user suddenly finds themselves lost, without a data
connection.

\section{Secure, Reliable, Scalable}
JEFRi has features to verify and secure any communication between different
JEFRi devices. This allows for very fine-grained control over what bits of data
are available to different users of the application from different devices.
JEFRi is built on solid practices, designed from the ground up to serve as a
framework to answer the challenges of the rapidly growing computing environment.
JEFRi is able to take advantage of both cloud computing and mobile platforms,
while still running on your desktop PC. With a test suite and full continual
integration, JEFRi continues to grow in features while maintaining a stable core
for future development.

\end{document}
